\textbf{Computational Fuzzy Extractors}

### 2013-03-25 trunk/CompFuzzExt.tex
Revised introduction

We will specifically focus on a one-round information reconciliation  mechanism called ``secure sketch,''  which, aside from its immediate application to authentication, is also
a crucial ingredient in the construction of fuzzy extractors~\cite{DBLP:journals/siamcomp/DodisORS08} (as well as in other contexts, e.g.,~\cite{Boyen05secureremote,dodisWichs2009,chandran2010privacy}).
It consists of  two algorithms: Sketch (used once) and Recover (used subsequently).  The Sketch ($\sketch$) algorithm takes an input $w$ and produces a sketch $s$.  This information allows
the Recover ($\rec$) algorithm to recover $w$ given $s$ and some value $w'$ that is close (according to some predefined metric, such as Hamming distance) to $w$. 
Crucially for security,  knowledge of $s$ should not reveal $w$; that is, $w$ should still have entropy  conditioned on $s$.  This feature is needed because $s$ is not secret: for example, in a single-user setting (where the user wants to recover the original reading $w$ from a subsequent reading $w'$), it would be stored in the clear; and in a key agreement application~\cite{Boyen05secureremote} (where two parties have $w$ and $w'$, respectively), it would be transmitted between the parties.
As defined in \cite{DBLP:journals/siamcomp/DodisORS08}, the entropy loss of a secure sketch is the difference in the entropy of $w$ and the entropy of $w$ conditioned on $s$.

Secret entropy of convenient sources is often at a premium; a goal of secure sketch constructions is to minimize the entropy loss, increasing the security of the resulting application. For example, in the application of secure sketches to fuzzy extractors (whose goal is to reliably produce a uniformly random secret key from a noisy input), the resulting secret key may be too short to be useful if the entropy loss is too high. However, because secure sketches are defined as information-theoretic objects, some entropy loss is inherent in any secure sketch construction, because $s$ contains enough information to recover $w$ from any value $w'$ that is close to it.

The minimum necessary  entropy loss has been quantified in past work.  Even if we don't require perfect correctness (i.e., allow $\rec$ to err with small probability~\cite[Section 8]{DBLP:journals/siamcomp/DodisORS08}), the entropy loss of a secure sketch that can reconcile $t$ errors with high probability is bounded by the redundancy of
the best code that can correct $t$ random errors with high probability~\cite[Proposition 8.2]{DBLP:journals/siamcomp/DodisORS08}
%(the argument is the same as in \cite[Lemma C.1]{DBLP:journals/siamcomp/DodisORS08}\lnote{probably we should double-check that this is correct}).

\paragraph {Our Contributions}
To see if entropy loss can be improved, we explore secure sketches with the relaxed secrecy requirement that is computational rather than information-theoretic.  
The natural relaxation of the secrecy requirement is to require HILL entropy~\cite{DBLP:journals/siamcomp/HastadILL99}~(namely, that the distribution of $w$ conditioned on $s$ be \emph{indistinguishable} from a high-min-entropy distribution).  Under this definition, using the paradigm for constructing fuzzy extractors from~\cite{DBLP:journals/siamcomp/DodisORS08} would yield a pseudorandom key.

On the negative side, we show defining sketches using HILL entropy is unlikely to be fruitful. We prove in Theorem~\ref{thm:impSketchArbitraryW} that the entropy loss of such secure sketches is subject to the same coding bounds as the ones that constrain information-theoretic secure sketches.
%Furthermore, we have (asymptotic)~constructions that meet these bounds.

On the positive side, we show that a different relaxation of the definition can in fact lead to a \emph{lossless} construction. The idea is to give up on reconstructing $w$ and, instead, to focus on producing a consistent string $\recout$  each time $\rec(w', s)$ is called.  This string $\recout$ is  required to have high HILL entropy even conditioned on $s$. Because this object \emph{conducts} the entropy from $w$ to $\recout$ and tolerates noise in its input, following \cite{CRVW02,KanukurthiR09}, we call it a \emph{computational fuzzy conductor}.

Given this relaxation, we are able to build \emph{lossless} computational fuzzy conductors, in which the HILL entropy of $\recout$ conditioned on $s$ is the same as the entropy of $w$ without conditioning on anything---i.e., before $\sketch$ is run (see Theorem~%s~%\ref{thm:lossless secure sketch} and
~\ref{thm:lossless secure conductor log})).  We demonstrate how to build such lossless conductors based on the Learning with Errors~(LWE) assumption due to Regev~\cite{regev2005LWE, regevLWEsurvey},  which says that decoding a random linear code is computationally difficult.

We show that this relaxation of the security requirement is  useful.  Specifically, we also define the notion of computational fuzzy extractors, in which the extracted key is required to be pseudorandom---i.e., to look random to any computationally bounded observer.
We show that our relaxed notion of secure sketches is sufficient to construct computational  fuzzy extractors via a simple and efficient construction, utilizing reconstructive extractors~\cite{barak-computational}.  The resulting fuzzy extractor produces a key that is as long as the entropy of $w$, minus only the loss necessary in any extractor construction.
\lnote{it would be good to find another application}

Both our lossless fuzzy conductor and the resulting fuzzy extractor are built via the code-offset construction~\cite{JW99},\cite[Section 5]{DBLP:journals/siamcomp/DodisORS08} already used in prior work.  To be able to apply the LWE assumption, we use a random linear code (unfortunately, our decoding algorithm can only reconcile a logarithmic number of differences).  We utilize the recent result of D\"{o}ttling and M\"{u}ller-Quade~\cite{dottling2012} that shows security of LWE when the error vector comes from the uniform distribution, with each coordinate ranging over a small interval.  The LWE error vector, in our case, is the input $w$ itself.  To utilize the result of~\cite{dottling2012}, we need to assume that the input comes from the uniform distribution. We are able to relax this limitation somewhat and allow $w$ to come from a symbol-fixing source~\cite{KZ07} (where each dimension is either uniform or fixed). This relaxation requires new results about the hardness of LWE when samples have a fixed error vector, which may be of independent interest~(\thref{thm:blockLWE}).

\section{Preliminaries}
\label{sec:preliminaries}
For a random variable $X = X_1||...|| X_n$ where each $X_i$ is over some alphabet $Z$, we denote by $X_{1,..., k} = X_1||...|| X_k$.  The {\em min-entropy} of $X$ is $\Hoo(X) = -\log(\max_x \Pr[X=x])$, 
%the {\em (worst-case) conditional min-entropy} of $X$ given $Y$ is  $\Hoo(X|Y) = -\log(\max_{x,y} \Pr[X=x|Y=y])$, and
and the {\em average (conditional)} min-entropy of $X$ given $Y$ is  $\Hav(X|Y) = -\log(\expe_{y\in Y} \max_{x} \Pr[X=x|Y=y])$~\cite[Section 2.4]{DBLP:journals/siamcomp/DodisORS08}.  
%The {\em collision probability} of  $X$ is 
%$\col(X) = \sum_{x} \Pr[X=x]^2$.
The {\em statistical distance} between random variables $X$ and $Y$ with the same domain is $\Delta(X,Y) = \frac12 \sum_x |\Pr[X=x] - \Pr[Y=x]|$. %We write $X \approx_{\epsilon} Y$ if $\Delta(X,Y) \leq \epsilon$, and when $\epsilon$ is negligible (in the appropriate parameter, as clear from the context) then we say $X$ and $Y$ are \emph{statistically close}.  
For a distinguisher $D$~(or a class of distinguishers $\mathcal{D}$) we write the \emph{computational distance} between $X$ and $Y$ as $\delta^D(X,Y) = \left| \expe[D(X)]-\expe[D(Y)]\right |$.  We denote by $\mathcal{D}_{s_{sec}}$ the class of randomized circuits which output a single bit and have size at most $s_{sec}$.
For a metric space $(\mathcal{M}, \dis)$, the \emph{(closed) ball of radius $t$ around $x$} is the set of all points within radius $t$, that is, $B_t(x) = \{y| \dis(x, y)\leq t\}$.  If the size of a ball in a metric space does not depend on $x$, we denote by $|B_t(\cdot)|$ the size of a ball of radius $t$.  For the Hamming metric over $Z^n$ the size of $B_t(\cdot) = \sum_{i=0}^t {n \choose t} (|Z|-1)^i $, for clarity we express this value as $B_t(\cdot)$.  $U_n$ denotes the uniformly  distributed random variable on $\{0,1\}^n$.
During the presentation, we denote bold letters to be vectors or matrix, capitalized letters to be random variables, and lower case letter to be the elements in the vector.

>>>>

One of the primary building blocks for authentications is reliable key derivation.  Fuzzy extractors~\cite{DBLP:journals/siamcomp/DodisORS08} achieve reliable key derivation from noisy sources.  (\bnote{Are these references for fuzzy extractors or secure sketches?} as well as in other contexts, e.g.,~\cite{Boyen05secureremote,dodisWichs2009,chandran2010privacy}).  The setting 
%We will specifically focus on a one-round information reconciliation  mechanism called ``secure sketch,''  which, aside from its immediate application to authentication, is also
%a crucial ingredient in the construction of fuzzy extractors~\cite{DBLP:journals/siamcomp/DodisORS08} (as well as in other contexts, e.g.,~\cite{Boyen05secureremote,dodisWichs2009,chandran2010privacy}).
consists of  two algorithms: Generate (used once) and Reproduce (used subsequently).  The Generate ($\gen$) algorithm takes an input $w$ and produces a key $r$ and a public value $p$.  This information allows
the Reproduce ($\rep$) algorithm to reproduce $r$ given $p$ and some value $w'$ that is close (according to some predefined metric, such as Hamming distance) to $w$. 
Crucially for security,  knowledge of $p$ should not reveal $r$; that is, $r$ should be uniformly distributed conditioned on $p$.  This feature is needed because $p$ is not secret: for example, in a single-user setting (where the user wants to recover the original reading $w$ from a subsequent reading $w'$), it would be stored in the clear; and in a key agreement application~\cite{Boyen05secureremote} (where two parties have $w$ and $w'$, respectively), it would be transmitted between the parties.

As defined in \cite{DBLP:journals/siamcomp/DodisORS08}, the entropy loss of a fuzzy extractor is the difference in the entropy of $w$ and the length of the key $r$.  A goal of fuzzy extractor constructions is to minimize the entropy loss, increasing the security of the resulting application.  The resulting secret key may be too short to be useful if the entropy loss is too high.  Traditionally, fuzzy extractors are defined as information-theoretic objects.  In the information-theoretic setting, some entropy loss is necessary as the value $p$ contains enough information to reproduce $r$ from any close value $w'$.

The main novel tool in constructing fuzzy extractors is the secure sketch.  The secure sketch produces a public value $s$ that allows recovery of $w$ from any close value $w'$.  The security requirement is that $w$ has high min-entropy conditioned on $s$.  The key $r$ is formed by applying a randomness extractor~\cite{nisan1993randomness} to $w$.
%Secret entropy of convenient sources is often at a premium; a goal of secure sketch constructions is to minimize the entropy loss, increasing the security of the resulting application. When used in a fuzzy extractor, the resulting secret key may be too short to be useful if the entropy loss is too high. 
The minimum entropy loss for secure sketches has been quantified in past work.  Even if we don't require perfect correctness (i.e., allow $\rep$ to err with small probability~\cite[Section 8]{DBLP:journals/siamcomp/DodisORS08}), the entropy loss of a secure sketch that can reconcile $t$ errors with high probability is bounded by the redundancy of
the best code that can correct $t$ random errors with high probability~\cite[Proposition 8.2]{DBLP:journals/siamcomp/DodisORS08}.  Known constructions of fuzzy extractors use a secure sketch suffer from this bound.
%(the argument is the same as in \cite[Lemma C.1]{DBLP:journals/siamcomp/DodisORS08}\lnote{probably we should double-check that this is correct}).

\paragraph {Our Contributions}
To see if entropy loss in fuzzy extractors can be improved, we explore fuzzy extractors and secure sketches with the relaxed secrecy requirement that is computational rather than information-theoretic.  Our overall goal is to produce a fuzzy extractor~(with small entropy loss), where the key $r$ is pseudorandom even conditioned on $p$.

On the negative side, we show that the paradigm of secure sketch and extraction from~\cite{DBLP:journals/siamcomp/DodisORS08} is unlikely to be fruitful.  We show that some entropy loss is necessary for secure sketches.
The most natural relaxation of the secrecy requirement is to require HILL entropy~\cite{DBLP:journals/siamcomp/HastadILL99}~(namely, that the distribution of $w$ conditioned on $s$ be \emph{indistinguishable} from a high-min-entropy distribution).  Under this definition, using the paradigm for constructing fuzzy extractors from~\cite{DBLP:journals/siamcomp/DodisORS08} would yield a pseudorandom key.  We prove in Theorem~\ref{thm:impSketchArbitraryW} that the entropy loss of such secure sketches is subject to the same coding bounds as the ones that constrain information-theoretic secure sketches.
%Furthermore, we have (asymptotic)~constructions that meet these bounds.

Another possible relaxation is to require that the value $w$ is still unpredictable conditioned on $s$.  Using an extractor with a reconstruction would still yield a pseudorandom key, see~\cite[Lemma 6]{DBLP:conf/eurocrypt/HsiaoLR07}.  We consider sketches that preserve the weaker notion of unpredictability for the Hamming metric.  In \thref{thm:imp of unp entropy}, we show  a significant loss in unpredictability is necessary for a secure sketch.  The entropy loss is proportional to the volume of the ball of radius $t$.

Both of these impossibility results arise because a sketch must function like an error-correcting code.  To avoid these impossibility results, we give up on reconstructing $w$ and focus directly on the entropy loss in fuzzy extractors.  
%to focus on producing a consistent string $\recout$  each time $\rec(w', s)$ is called.  This string $\recout$ is  required to have high HILL entropy even conditioned on $s$. Because this object \emph{conducts} the entropy from $w$ to $\recout$ and tolerates noise in its input, following \cite{CRVW02,KanukurthiR09}, we call it a \emph{computational fuzzy conductor}.
We build \emph{lossless} computational fuzzy extractors, in which $r$ is pseudorandom conditioned on $p$ and $r$ is as long as the entropy of $w$.  % without conditioning on anything---i.e., before $\gen$ is run (see~\thref{thm:lossless secure conductor log})).  
We demonstrate how to build such lossless fuzzy extractors~(\consref{cons:informal construction}) based on the Learning with Errors~(LWE) assumption due to Regev~\cite{regev2005LWE, regevLWEsurvey},  which says that decoding a random linear code is computationally difficult.

%We show that this relaxation of the security requirement is  useful.  Specifically, we also define the notion of computational fuzzy extractors, in which the extracted key is required to be pseudorandom---i.e., to look random to any computationally bounded observer.
%We show that our relaxed notion of secure sketches is sufficient to construct computational  fuzzy extractors via a simple and efficient construction, utilizing reconstructive extractors~\cite{barak-computational}.  The resulting fuzzy extractor produces a key that is as long as the entropy of $w$, minus only the loss necessary in any extractor construction.
%\lnote{it would be good to find another application}

Our lossless fuzzy extractor is built via the code-offset construction~\cite{JW99},\cite[Section 5]{DBLP:journals/siamcomp/DodisORS08} already used in prior work.  To be able to apply the LWE assumption, we use a random linear code (unfortunately, our decoding algorithm can only reconcile a logarithmic number of differences).  We utilize the recent result of D\"{o}ttling and M\"{u}ller-Quade~\cite{dottling2012} that shows security of LWE when the error vector comes from the uniform distribution, with each coordinate ranging over a small interval.  The LWE error vector, in our case, is the input $w$ itself.  To utilize the result of~\cite{dottling2012}, we need to assume that the input comes from the uniform distribution. We are able to relax this limitation somewhat and allow $w$ to come from a symbol-fixing source~\cite{KZ07} (where each dimension is either uniform or fixed). This relaxation requires new results about the hardness of LWE when samples have a fixed~(and adversarially known) error vector, which may be of independent interest~(\thref{thm:blockLWE}).

%Authentication generally requires secrets.  Unfortunately, many useful  sources that contain sufficient entropy for a secret  are also noisy, and provide similar, but not identical secret values at each invocation (examples of such sources include biometrics~\cite{daugman2004}, human memory~\cite{zviran1993comparison}, pictorial passwords~\cite{brostoff2000passfaces}, measurements of capacitance~\cite{tuyls2006puf}, timing~\cite{suh2007physical}, motion~\cite{castelluccia2005shake},  quantum information~\cite{bennett1988privacy} etc.).  Information reconciliation protocols~\cite{bennett1988privacy} remove the noise without revealing the secrets.
%We will specifically focus on a one-round information reconciliation  mechanism called ``secure sketch,''  which, aside from its immediate application to authentication, is also
%a crucial ingredient in the construction of fuzzy extractors~\cite{DBLP:journals/siamcomp/DodisORS08} (as well as in other contexts, e.g.,~\cite{Boyen05secureremote,dodisWichs2009,chandran2010privacy}).
%It consists of  two algorithms: Sketch (used once) and Recover (used subsequently).  The Sketch ($\sketch$) algorithm takes an input $w$ and produces a sketch $s$.  This information allows
%the Recover ($\rec$) algorithm to recover $w$ given $s$ and some value $w'$ that is close (according to some predefined metric, such as Hamming distance) to $w$. 
%Crucially for security,  knowledge of $s$ should not reveal $w$; that is, $w$ should still have entropy  conditioned on $s$.  This feature is needed because $s$ is not secret: for example, in a single-user setting (where the user wants to recover the original reading $w$ from a subsequent reading $w'$), it would be stored in the clear; and in a key agreement application~\cite{Boyen05secureremote} (where two parties have $w$ and $w'$, respectively), it would be transmitted between the parties.
%As defined in \cite{DBLP:journals/siamcomp/DodisORS08}, the entropy loss of a secure sketch is the difference in the entropy of $w$ and the entropy of $w$ conditioned on $s$.  
%Secret entropy of convenient sources is often at a premium; a goal of secure sketch constructions is to minimize the entropy loss, increasing the security of the resulting application. For example, in the application of secure sketches to fuzzy extractors (whose goal is to reliably produce a uniformly random secret key from a noisy input), the resulting secret key may be too short to be useful if the entropy loss is too high. However, because secure sketches are defined as information-theoretic objects, some entropy loss is inherent in any secure sketch construction, because $s$ contains enough information to recover $w$ from any value $w'$ that is close to it.  
%The minimum necessary  entropy loss has been quantified in past work.  Even if we don't require perfect correctness (i.e., allow $\rec$ to err with small probability~\cite[Section 8]{DBLP:journals/siamcomp/DodisORS08}), the entropy loss of a secure sketch that can reconcile $t$ errors with high probability is bounded by the redundancy of
%the best code that can correct $t$ random errors with high probability~\cite[Proposition 8.2]{DBLP:journals/siamcomp/DodisORS08}
%%(the argument is the same as in \cite[Lemma C.1]{DBLP:journals/siamcomp/DodisORS08}\lnote{probably we should double-check that this is correct}).
%\paragraph {Our Contributions}
%To see if entropy loss can be improved, we explore secure sketches with the relaxed secrecy requirement that is computational rather than information-theoretic.  
%The natural relaxation of the secrecy requirement is to require HILL entropy~\cite{DBLP:journals/siamcomp/HastadILL99}~(namely, that the distribution of $w$ conditioned on $s$ be \emph{indistinguishable} from a high-min-entropy distribution).  Under this definition, using the paradigm for constructing fuzzy extractors from~\cite{DBLP:journals/siamcomp/DodisORS08} would yield a pseudorandom key.  
%On the negative side, we show defining sketches using HILL entropy is unlikely to be fruitful. We prove in Theorem~\ref{thm:impSketchArbitraryW} that the entropy loss of such secure sketches is subject to the same coding bounds as the ones that constrain information-theoretic secure sketches.
%%Furthermore, we have (asymptotic)~constructions that meet these bounds.
%On the positive side, we show that a different relaxation of the definition can in fact lead to a \emph{lossless} construction. The idea is to give up on reconstructing $w$ and, instead, to focus on producing a consistent string $\recout$  each time $\rec(w', s)$ is called.  This string $\recout$ is  required to have high HILL entropy even conditioned on $s$. Because this object \emph{conducts} the entropy from $w$ to $\recout$ and tolerates noise in its input, following \cite{CRVW02,KanukurthiR09}, we call it a \emph{computational fuzzy conductor}.
%Given this relaxation, we are able to build \emph{lossless} computational fuzzy conductors, in which the HILL entropy of $\recout$ conditioned on $s$ is the same as the entropy of $w$ without conditioning on anything---i.e., before $\sketch$ is run (see Theorem~%s~%\ref{thm:lossless secure sketch} and
%~\ref{thm:lossless secure conductor log})).  We demonstrate how to build such lossless conductors based on the Learning with Errors~(LWE) assumption due to Regev~\cite{regev2005LWE, regevLWEsurvey},  which says that decoding a random linear code is computationally difficult. 
%We show that this relaxation of the security requirement is  useful.  Specifically, we also define the notion of computational fuzzy extractors, in which the extracted key is required to be pseudorandom---i.e., to look random to any computationally bounded observer.
%We show that our relaxed notion of secure sketches is sufficient to construct computational  fuzzy extractors via a simple and efficient construction, utilizing reconstructive extractors~\cite{barak-computational}.  The resulting fuzzy extractor produces a key that is as long as the entropy of $w$, minus only the loss necessary in any extractor construction.
%\lnote{it would be good to find another application} 
%Both our lossless fuzzy conductor and the resulting fuzzy extractor are built via the code-offset construction~\cite{JW99},\cite[Section 5]{DBLP:journals/siamcomp/DodisORS08} already used in prior work.  To be able to apply the LWE assumption, we use a random linear code (unfortunately, our decoding algorithm can only reconcile a logarithmic number of differences).  We utilize the recent result of D\"{o}ttling and M\"{u}ller-Quade~\cite{dottling2012} that shows security of LWE when the error vector comes from the uniform distribution, with each coordinate ranging over a small interval.  The LWE error vector, in our case, is the input $w$ itself.  To utilize the result of~\cite{dottling2012}, we need to assume that the input comes from the uniform distribution. We are able to relax this limitation somewhat and allow $w$ to come from a symbol-fixing source~\cite{KZ07} (where each dimension is either uniform or fixed). This relaxation requires new results about the hardness of LWE when samples have a fixed error vector, which may be of independent interest~(\thref{thm:blockLWE}).
\section{Preliminaries}
\label{sec:preliminaries}
For a random variable $X = X_1||...|| X_n$ where each $X_i$ is over some alphabet $Z$, we denote by $X_{1,..., k} = X_1||...|| X_k$.  The {\em min-entropy} of $X$ is $\Hoo(X) = -\log(\max_x \Pr[X=x])$, 
%the {\em (worst-case) conditional min-entropy} of $X$ given $Y$ is  $\Hoo(X|Y) = -\log(\max_{x,y} \Pr[X=x|Y=y])$, and
and the {\em average (conditional)} min-entropy of $X$ given $Y$ is  $\Hav(X|Y) = -\log(\expe_{y\in Y} \max_{x} \Pr[X=x|Y=y])$~\cite[Section 2.4]{DBLP:journals/siamcomp/DodisORS08}.  
%The {\em collision probability} of  $X$ is 
%$\col(X) = \sum_{x} \Pr[X=x]^2$.
The {\em statistical distance} between random variables $X$ and $Y$ with the same domain is $\Delta(X,Y) = \frac12 \sum_x |\Pr[X=x] - \Pr[Y=x]|$. %We write $X \approx_{\epsilon} Y$ if $\Delta(X,Y) \leq \epsilon$, and when $\epsilon$ is negligible (in the appropriate parameter, as clear from the context) then we say $X$ and $Y$ are \emph{statistically close}.  
For a distinguisher $D$~(or a class of distinguishers $\mathcal{D}$) we write the \emph{computational distance} between $X$ and $Y$ as $\delta^D(X,Y) = \left| \expe[D(X)]-\expe[D(Y)]\right |$.  We denote by $\mathcal{D}_{s_{sec}}$ the class of randomized circuits which output a single bit and have size at most $s_{sec}$.
For a metric space $(\mathcal{M}, \dis)$, the \emph{(closed) ball of radius $t$ around $x$} is the set of all points within radius $t$, that is, $B_t(x) = \{y| \dis(x, y)\leq t\}$.  If the size of a ball in a metric space does not depend on $x$, we denote by $|B_t(\cdot)|$ the size of a ball of radius $t$.  For the Hamming metric over $Z^n$ the size of $B_t(\cdot) = \sum_{i=0}^t {n \choose t} (|Z|-1)^i $, for clarity we express this value as $B_t(\cdot)$.  $U_n$ denotes the uniformly  distributed random variable on $\{0,1\}^n$.
During the presentation, we denote bold letters to be vectors or matrix, capitalized letters to be random variables, and lower case letter to be the elements in the vector.
